---
title: "project2"
author: "Xiaomeng Liu"
date: "2023-10-04"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# overall goal\

Our goal with this project is to create a vignette about contacting an API using functions you’ve created to
query, parse, and return well-structured data. You’ll then use your functions to obtain data from the API and
do some exploratory data analysis. This is an individual project and your work will be done in a github repo.

# Setting up your Github Repo


# Vignette Content Details
# requirement

## write functions to contact your chosen API and return 

```{r}
library(tidyverse) ; library(httr) ; library(jsonlite)
```

## Fucntions to interact with the API

### create a function for basic url generation 
```{r}
# create a function for basic url generation
pu<-function(endpoint){ 
    # param1 is mod and 2 is end
    ori_url="https://api.nasa.gov/neo/rest/v1"
  apiurl<-paste(ori_url,endpoint,sep = "/")
  
  lastchar<-substr(apiurl,nchar(apiurl),nchar(apiurl))
  if(lastchar=="?"|lastchar=="/"){
    apiurl<-substr(apiurl,1,nchar(apiurl)-1)
  }
  apiurl
  }
```

### create a function to ineract with the first endpoint 
```{r}
# create a function to interact with one end point
mo<-function(apiurl,mod){
  dat<-paste0("start_date=",mod)
  a<-paste(apiurl,dat,sep = "?")
  a<-paste(a,"api_key=yUhdPcxBc0yGQ0IeWdf1DFPHW31gJLtDL8RAiL0b",sep = "&")
  a
}
```

```{r}
# create a function to draw it from API
da<-function(endpoint,mod){
  outputAPI<-fromJSON(mo(pu(endpoint),mod))
  outputAPI
}
```



```{r}
# function for mod2
mo2<-function(outputAPI,mod){
  outputAPI<-outputAPI$near_earth_objects[noquote(mod)]
outputAPI
  
}

# function for mod3
mo3<-function(outputAPI,mod){
  ifelse(is.null(mod),print("missing value"),
         output<-outputAPI[[1]]$close_approach_data)
  output<-data.frame(matrix(unlist(output),nrow=length(output),byrow = TRUE))
  colnames(output)<-c("close_approach_date", "close_approach_date_full", "epoch_date_close_approach", "relative_velocity.kilometers_per_second","relative_velocity.kilometers_per_hour", "relative_velocity.miles_per_hour", "miss_distance.astronomical",
                      "miss_distance.lunar", "miss_distance.kilometers", "miss_distance.miles", "orbiting_body")
  output$epoch_date_close_approach<-as.numeric(output$epoch_date_close_approach)
  ifelse(mod<max(output$epoch_date_close_approach),
  output<-output%>% filter(epoch_date_close_approach > mod),
  print("too large for the epoch date close approach"))
  # a warning message will be shown and the filter will not be applied
  output # output the result anyway
}
```

```{r}
# define the function for all the end point and mods
# mod1 and mod4 are for the start_data which could be empty as default
# mod2 would be `is_potentially_hazardous_asteroid`
# mod3 would be `close_approach_data$epoch_date_close_approach`
finalcal<-function(end1,mod1="none",mod2,mod3){
  # for end1
  outputAPI<-da(endpoint=end1,mod1)
  outputAPI<-mo2(outputAPI,mod2)
  output<-mo3(outputAPI,mod3)
  output
}
```

```{r}
# do a test
finalcal(end1 = "feed",mod1 = "2015-09-07",
         mod2 = "2015-09-08",mod3 = 1.7e12)
```

```{r}
#outputAPI<-da("feed","start_date=2015-09-07")
#outputAPI<-mo2(outputAPI,mod="2015-09-08")
#output<-mo3(outputAPI,mod=1.4e12)
```

```{r}
# create a function for second endpoint lookup
da2<-function(endpoint,mod){
  mo_new<-function(apiurl,mod){
  a<-paste(apiurl,mod,sep = "/")
  a<-paste(a,"api_key=yUhdPcxBc0yGQ0IeWdf1DFPHW31gJLtDL8RAiL0b",sep = "?")
  a
  }
  outputAPI<-fromJSON(mo_new(pu(endpoint),mod))
  output<-outputAPI$close_approach_data
  output<-data.frame(matrix(unlist(output),ncol = 11,byrow = FALSE))
  colnames(output)<-c("close_approach_date", "close_approach_date_full", "epoch_date_close_approach", "relative_velocity.kilometers_per_second","relative_velocity.kilometers_per_hour", "relative_velocity.miles_per_hour", "miss_distance.astronomical",
                      "miss_distance.lunar", "miss_distance.kilometers", "miss_distance.miles", "orbiting_body")
  output
}
```

```{r}
# create a function for both endpoints
both<-function(end1,mod1,mod2,mod3,end2,mod4="none"){
  n1<-finalcal(end1,mod1 = mod1,mod2 = mod2,mod3 = mod3)
  n2<-da2(endpoint = end2,mod = mod4)
  list(n1,n2)
}
```

find the ID using this tool for endpoint2: https://ssd.jpl.nasa.gov/tools/sbdb_query.html#!#results

```{r}
# test the both function
result<-both(end1 = "feed",mod1 = "2015-09-07",
         mod2 = "2015-09-08",mod3 = 1.7e12,
     end2 = "neo",mod4 = 3542517
     )
result
```
```{r}
# run the both function again for data used to be compared
# the question would be finding the difference between different time and different aest
result2<-both(end1 = "feed",mod1 = "2023-09-07",
         mod2 = "2023-09-08",mod3 = 1.7e12,
     end2 = "neo",mod4 = 3542519
     )
result2
```

```{r}
# since we want to compare them separately, I would combine them into two tables
summary(result[[1]])
```
```{r}
miss_distance<-result[[1]]%>%
  summarize("min"=min(miss_distance.astronomical),
            "max"=max(miss_distance.astronomical))
knitr::kable(miss_distance,caption="example")

```

```{r}
data_2015<-result[[1]]
data_2023<-result2[[1]]
num_1<-result[[2]]
num_2<-result2[[2]]
```

```{r}
data_together<-rbind(data_2015,data_2023)
```

```{r}
str(data_together)
```
```{r}
# plot1
plot1<-ggplot(data_together,aes(x=close_approach_date))
plot1+geom_boxplot(aes(y=relative_velocity.kilometers_per_second))
plot1+geom_boxplot(aes(y=as.numeric(miss_distance.lunar)))
```

```{r}
num_1$epoch_date_close_approach<-as.numeric(num_1$epoch_date_close_approach)
num_1$relative_velocity.kilometers_per_second<-as.numeric(num_1$relative_velocity.kilometers_per_second)
# plot 2
plot2<-ggplot(num_1,aes(x=close_approach_date,y=epoch_date_close_approach))
plot2+geom_col()
```