---
title: "project2"
author: "Xiaomeng Liu"
date: "2023-10-04"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# overall goal\

Our goal with this project is to create a vignette about contacting an API using functions you’ve created to
query, parse, and return well-structured data. You’ll then use your functions to obtain data from the API and
do some exploratory data analysis. This is an individual project and your work will be done in a github repo.

# Setting up your Github Repo


# Vignette Content Details

## write functions to contact your chosen API and return 

```{r}
library(tidyverse) ; library(httr) ; library(jsonlite)
```

## Fucntions to interact with the API

```{r}
# create a function for ends input
pu<-function(endpoint){ 
    # param1 is mod and 2 is end
    ori_url="https://api.nasa.gov/neo/rest/v1"
  apiurl<-paste(ori_url,endpoint,sep = "/")
  
  lastchar<-substr(apiurl,nchar(apiurl),nchar(apiurl))
  if(lastchar=="?"|lastchar=="/"){
    apiurl<-substr(apiurl,1,nchar(apiurl)-1)
  }
  apiurl
  }
```

```{r}
# create a function to interact with one end point
mo<-function(apiurl,mod){
  dat<-paste0("start_date=",mod)
  a<-paste(apiurl,dat,sep = "?")
  a<-paste(a,"api_key=yUhdPcxBc0yGQ0IeWdf1DFPHW31gJLtDL8RAiL0b",sep = "&")
  a
}
```

```{r}
# create a function to draw it from API
da<-function(endpoint,mod){
  outputAPI<-fromJSON(mo(pu(endpoint),mod))
  outputAPI
}
```



```{r}
# function for mod2
mo2<-function(outputAPI,mod){
#if(mod==`2023-10-08`){
  outputAPI<-outputAPI$near_earth_objects[noquote(mod)]
#} 

outputAPI
  
}

# function for mod3
mo3<-function(outputAPI,mod){
  ifelse(is.null(mod),print("missing value"),
         output<-outputAPI[[1]]$close_approach_data)
  output<-data.frame(matrix(unlist(output),nrow=length(output),byrow = TRUE))
  colnames(output)<-c("close_approach_date", "close_approach_date_full", "epoch_date_close_approach", "relative_velocity.kilometers_per_second","relative_velocity.kilometers_per_hour", "relative_velocity.miles_per_hour", "miss_distance.astronomical",
                      "miss_distance.lunar", "miss_distance.kilometers", "miss_distance.miles", "orbiting_body")
  output$epoch_date_close_approach<-as.numeric(output$epoch_date_close_approach)
  ifelse(mod<max(output$epoch_date_close_approach),
  output<-output%>% filter(epoch_date_close_approach > mod),
  print("too large for the epoch date close approach"))
  # a warning message will be shown and the filter will not be applied
  output # output the result anyway
}
```

```{r}
# define the function for all the end point and mods
# mod1 and mod4 are for the start_data which could be empty as default
# mod2 would be `is_potentially_hazardous_asteroid`
# mod3 would be `close_approach_data$epoch_date_close_approach`
finalcal<-function(end1,mod1="none",mod2,mod3){
  # for end1
  outputAPI<-da(endpoint=end1,mod1)
  outputAPI<-mo2(outputAPI,mod2)
  output<-mo3(outputAPI,mod3)
  output
}
```

```{r}
# do a test
finalcal(end1 = "feed",mod1 = "2015-09-07",
         mod2 = "2015-09-08",mod3 = 1.7e12)
```

```{r}
outputAPI<-da("feed","start_date=2015-09-07")
outputAPI<-mo2(outputAPI,mod="2015-09-08")
output<-mo3(outputAPI,mod=1.4e12)
```


